a and generators

### 2022-05-10 documentos/pce.tex
single-qubit
single-qubit en todos

% }}}
\subsection{Structure of PCE maps} % {{{
\label{sec:single:qubit}
% \todo{David escribe una primera version y luego Carlos itera}
% Now that we have discussed the main features of quantum channels
% we are going to introduce the Pauli component erasing maps.
% We shall start exploring the single qubit scenario 
% and continue with the $N$-qubit case. 
% 
% \afnote{
We have discussed the main features of quantum channels, now
we turn our attention to introduce the Pauli component erasing maps. We start
by exploring the single qubit scenario and then we treat the $N$-qubit case.
% }

The most general single qubit density matrix can be written as
\begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
\rho=\frac12 {\sum_{\alpha=0}^3 r_\alpha \sigma_\alpha},
\end{equation}
% \afnote{Creo que este párrafo queda mejor así: where
% $\{\sigma_{0},\sigma_{1},\sigma_{2},\sigma_{3}\}=\{\one,\sigma_{x},\sigma_{y},\sigma_{z}\}$,
% with $\one$ the $2\times 2$ identity matrix and $\sigma_{j}$ the usual Pauli
% matrices. Note that normalization requires that $r_0=1$. On the other hand, the
% remaining $r_{j}$ coefficients constitute the so called Bloch vector
% $\vec{r}=(r_{x},r_{y},r_{z})$. Consider the map that projects each component in
% the following way:}
% 
with $\sigma_0= \one$, and $\sigma_{1,2,3}$ the usual Pauli matrices.  
Normalization requires that $r_0=1$ and the remaining $r_{1,2,3}$ 
form the Bloch vector. 
Consider the map that projects
each component 
in the following way,
% 
\begin{equation}
r_\alpha \mapsto \tau_\alpha r_\alpha
\label{EqPCE}
\end{equation}
where $\tau_\alpha$ is either $0$ or $1$ (trace preserving requires
that $\tau_0=1$). 
From now on we refer to any operation like that described in
\eref{EqPCE}, as single-aco aqubit \textit{Pauli component erasing} (\pce{}) maps.
Not every such operation  is a quantum channel;
for example, collapsing the entire Bloch ball to a disk on the $xy$
plane ($\tau_{1}=\tau_{2}=1$ and  $\tau_3=0$) 
leads
to a violation of the CP conditions. Indeed a direct evaluation
of such conditions yield~\cite{zimansbook,davalosdivisibility}, 
% 
% 
% is not a quantum channel
% \afnote{(Aquí podemos colocar en vez de ``is not a quantum channel'', ``leads
% to a violation of the CP conditions, indeed a direct evaluation
% \cite{zimansbook,davalosdivisibility}'')}. A direct evaluation of the complete
% positivity conditions~\cite{zimansbook,davalosdivisibility}, 
\begin{align}
1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
1+\tau_\alpha-\tau_\beta-\tau_\gamma &\geq 0 \ \ \forall ~\alpha\neq \beta \neq \gamma,
\end{align}
where trace-preserving is already imposed, \janote{and} show that five out of
the eight 1-qubit PCE maps are quantum channels.
% 
% show that only five
% operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
These operations are the identity map, the completely depolarizing channel
($\rho \mapsto \one/2$); and bit, phase, and the bit-phase flip (with flip probability of $1/2$)
channels~\cite{chuangbook}, 
% These operations are by definition diagonal in the
% Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat
% \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$
and can be pictured using one column tables showing the positions of $0$s and
$1$s, see~\fref{fig:one:qubit:examples}.

@@

in the diagrams that represent the single qubit \pce{} maps,

@@

% representation of a single qubit \pce{} map.

@@

Similarly to the single qubit case, the density matrix $\rho$ of a system of $N$
qubits can be written using Pauli strings in the following way, 
\begin{equation}\label{eq:N_qubits_rho}
	\rho =\frac{1}{2^N} 
            \sum_\valpha r_\valpha \sigma_\valpha,
%             \quad r_{\vec0}=1,
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
%             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
	%r_{0,\ldots,0}=1,
\end{equation}
so $r_{\vec \alpha}=\langle \sigma_{\vec \alpha} \rangle=\tr \left(\rho
\sigma_{\vec \alpha} \right)$ is the coefficient corresponding to the expansion
of the density matrix in the normalized basis of Pauli strings.  Again,
normalization of the state requires that $r_{\vec 0}=1$.
We shall refer to $\paulicomponents$ as the {\it Pauli components} of the density
matrix of a system of qubits.

>>>>

% }}}
\subsection{Structure of PCE maps} % {{{
\label{sec:single:qubit}
% \todo{David escribe una primera version y luego Carlos itera}
% Now that we have discussed the main features of quantum channels
% we are going to introduce the Pauli component erasing maps.
% We shall start exploring the single qubit scenario 
% and continue with the $N$-qubit case. 
% 
% \afnote{
We have discussed the main features of quantum channels, now
we turn our attention to introduce the Pauli component erasing maps. We start
by exploring the single-qubit scenario and then we treat the $N$-qubit case.
% }

The most general single-qubit density matrix can be written as
\begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
\rho=\frac12 {\sum_{\alpha=0}^3 r_\alpha \sigma_\alpha},
\end{equation}
% \afnote{Creo que este párrafo queda mejor así: where
% $\{\sigma_{0},\sigma_{1},\sigma_{2},\sigma_{3}\}=\{\one,\sigma_{x},\sigma_{y},\sigma_{z}\}$,
% with $\one$ the $2\times 2$ identity matrix and $\sigma_{j}$ the usual Pauli
% matrices. Note that normalization requires that $r_0=1$. On the other hand, the
% remaining $r_{j}$ coefficients constitute the so called Bloch vector
% $\vec{r}=(r_{x},r_{y},r_{z})$. Consider the map that projects each component in
% the following way:}
% 
with $\sigma_0= \one$, and $\sigma_{1,2,3}$ the usual Pauli matrices.  
Normalization requires that $r_0=1$ and the remaining $r_{1,2,3}$ 
form the Bloch vector. 
Consider the map that projects
each component 
in the following way,
% 
\begin{equation}
r_\alpha \mapsto \tau_\alpha r_\alpha
\label{EqPCE}
\end{equation}
where $\tau_\alpha$ is either $0$ or $1$ (trace preserving requires
that $\tau_0=1$). 
From now on we refer to any operation like that described in
\eref{EqPCE}, as single-aco aqubit \textit{Pauli component erasing} (\pce{}) maps.
Not every such operation  is a quantum channel;
for example, collapsing the entire Bloch ball to a disk on the $xy$
plane ($\tau_{1}=\tau_{2}=1$ and  $\tau_3=0$) 
leads
to a violation of the CP conditions. Indeed a direct evaluation
of such conditions yield~\cite{zimansbook,davalosdivisibility}, 
% 
% 
% is not a quantum channel
% \afnote{(Aquí podemos colocar en vez de ``is not a quantum channel'', ``leads
% to a violation of the CP conditions, indeed a direct evaluation
% \cite{zimansbook,davalosdivisibility}'')}. A direct evaluation of the complete
% positivity conditions~\cite{zimansbook,davalosdivisibility}, 
\begin{align}
1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
1+\tau_\alpha-\tau_\beta-\tau_\gamma &\geq 0 \ \ \forall ~\alpha\neq \beta \neq \gamma,
\end{align}
where trace-preserving is already imposed, \janote{and} show that five out of
the eight 1-qubit PCE maps are quantum channels.
% 
% show that only five
% operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
These operations are the identity map, the completely depolarizing channel
($\rho \mapsto \one/2$); and bit, phase, and the bit-phase flip (with flip probability of $1/2$)
channels~\cite{chuangbook}, 
% These operations are by definition diagonal in the
% Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat
% \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$
and can be pictured using one column tables showing the positions of $0$s and
$1$s, see~\fref{fig:one:qubit:examples}.

@@

in the diagrams that represent the single-qubit \pce{} maps,

@@

% representation of a single-qubit \pce{} map.

@@

Similarly to the single-qubit case, the density matrix $\rho$ of a system of $N$
qubits can be written using Pauli strings in the following way, 
\begin{equation}\label{eq:N_qubits_rho}
	\rho =\frac{1}{2^N} 
            \sum_\valpha r_\valpha \sigma_\valpha,
%             \quad r_{\vec0}=1,
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
%             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
	%r_{0,\ldots,0}=1,
\end{equation}
so $r_{\vec \alpha}=\langle \sigma_{\vec \alpha} \rangle=\tr \left(\rho
\sigma_{\vec \alpha} \right)$ is the coefficient corresponding to the expansion
of the density matrix in the normalized basis of Pauli strings.  Again,
normalization of the state requires that $r_{\vec 0}=1$.
We shall refer to $\paulicomponents$ as the {\it Pauli components} of the density
matrix of a system of qubits.
